Generate Real and Pseudo-absences

### 2023-12-17 scripts/generate_absences.Rmd
preparing lda fn for PA removals

out <- lapply(records, absence_drawer, bg_abs = 0.15)
out <- bind_rows(out)
st_write(out, '../data/raw/occurrence/absences.shp')
```

>>>>

out <- lapply(records, absence_drawer, bg_abs = 0.15)
out <- bind_rows(out)
st_write(out, '../data/raw/occurrence/absences.shp')

ab <- st_read('../data/raw/occurrence/absences.shp')
```

```{r remove pseudo absences in presence niche space}

library(klaR)
library(psych)
library(MASS)
library(ggord)

# to determine whether the pseudo-random absences are in areas which are possibly suitable
# we will use linear discrimination analysis (LDA). LDA uses a label, e.g. presence or absence
# as a response which can be predicted by independent variables which are linear combinations 
# generated by reducing a higher dimensional data set to two axis. LDA will be trained on a data set 
# of AIM Plots with true absences, geographic absences, and a subset of the Psuedo-absences. 
# Only up to 750:750 presences:absences per species will be used to train a model. The maximum
# imbalance between presences to absences will be 750:250, or 3:1.

lda_PA_dropper <- function(x, path){

  pres <- x[x$Occurrence == 'Presence', ]
  real_abs <- x[x$PtType == 'AIM Absence', ]
  geo_abs <- x[x$PtType == 'Background Absence', ]
  model_abs <- dplyr::bind_rows(real_abs, geo_abs)

  if(nrow(pres) > 750){pres <- sample(1:nrow(pres), size = 750, replace = FALSE)}
  if(nrow(model_abs) > 750){model_abs <- sample(1:nrow(model_abs), size = 750, replace = FALSE)}

  training <- dplyr::bind_rows(pres, model_abs)
  testing <- filter(x, ! ID %in% training$ID)

  LDA_model <- lda(Species ~ ., training) # generate model

  # Here the accuracy of the trained modeled is evaluated
  LDA_train <- predict(LDA_model, training)$class
  tab <- table(Predicted = LDA_train, Actual = training$Species)
  ACC_train <- sum(diag(tab))/sum(tab) # accuracy

  # Here the model is applied to unclassified data
  LDA_test <- predict(LDA_model, testing)$class
  testing$Predicted <- predict(LDA_model, testing)$class
  tab1 <- table(Predicted = LDA_test, Actual = testing$Species)
  ACC_test <- sum(diag(tab1))/sum(tab1) # accuracy.

  n_pres <- nrow(x[x$Occurrence == 'Presence',])
  n_abs <- nrow(x[x$Occurrence != 'Presence',])
  conflicted <- filter(testing, Predicted != Occurrence) 
  max_rcs_2_remove <- n_pres - n_ab

  if(nrow(conflicted) == 0){ # identify the records we want to remove from the test data. 
    removals <- NA
  } else if (nrow(conflicted) > max_rcs_2_remove) {
    removals <- conflicted[sample(1:nrow(conflicted), size = max_rcs_2_remove, replace = FALSE), 'ID']
  } else {
    removals <- conflicted$ID
  }

  if(!is.na(removals)){data_out <- dplyr::filter(x, ! ID %in% removals)}

  metrics <- data.frame(
    Presences = nrow(pres), 
    Absences = nrow(model_abs),
    ACC_train = ACC_train, 
    ACC_test = ACC_test,
    Flagged = nrow(conflicted)
    )

  species <- x$Taxon[1]
  message(length(removals), ' records dropped from ', species)
  species <- gsub(' ', '_', species)
  write.csv(metrics, file.path(path, paste0(species, '.csv')), row.names = FALSE)

  return(data_out)
}

```
